### 2020-10-06 hpp.tex
Write state of the art and start experimental results.

\section{Introduction}
This \paper describes a software platform taylored for manipulation planning in
robotics.
The main contributions are:
\begin{itemize}
\item an original and general implementation of prehensile manipulation based on
  non-linear constraints,
\item an original solver for non-linear constraints that is able to handle implicit and explicit constraints,
\item a manipulation planning algorithm that is able to tackle a great variety of manipulation planning problems,
\item an open-source software suite that implements all the above following state of the art development tools and methods.
\end{itemize}

The \paper is organized as follows.
\Partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint. \Partie~\ref{sec:constraints} introduces non linear constraints and solvers that are at the core of manipulation problem definition. \Partie~\ref{sec:manipulation-problem} defines the problem of prehensile manipulation in the general setting. \Partie~\ref{sec:manipulation-planning} provides a general algorithm that solves manipulation planning problems. Finally, \partie~\ref{sec:humanoid-path-planner} is devoted to a software platform implementing the notions introduced in the previous sections.

Each \partie is implemented by one or several software packages. Instead of providing formulas for some values that need to be computed, we provide a link to the C++ implementation.

@@

Motion planning has given rise to a lot of research work for the past decades.
The problem consists in finding a collision-free path for a given system in
an environment populated with obstacles. The field covers a large variety of
different applications ranging from navigation for autonomous vehicles in
partially known environments~\cite{DarpaUrbanChallenge} to path planning for
deformable objects~\cite{LamKav2001},\\
\cite{RouFerTai2020} and many other applications like coverage path planning~\cite{coverage2001},\\ \cite{coverage2013}, or pursuit evasion planning~\cite{pursuit-evasion1999}.

Planning motions for high dimensional robots like humanoid robots or multi-arm
system has been proven to be of a very high complexity~\cite{SchSha1983},\\ \cite{Canny1983}. Starting in the 1990's random sampling methods have been proposed
to solve the problem, trading the completeness property against efficiency in
solving problem in high dimensional configuration spaces\\ \cite{KSLO1996}, \cite{HsuLatMot99}, \cite{KufLav00}. Since then, asymptotically optimal random sampling algorithms have
been proposed~\cite{KarFra2011}.

@@

Some systems are subject to non-linear constraints. These non-linear constraints
define a sub-manifold of the configuration the robot must stay on. For instance,
legged robots that must keep contact with the ground and enforce quasi-static
equilibrium, or multi-arm systems grasping a same object are subject to this
type of constraints. As the volume of the constrained manifold is usually equal
to zero, sampling random configurations satisfying the constraints is an event
of zero probability. To sample configurations on the constrained manifold,
\cite{DalNakLamLau2009} and \cite{berenson2011} projects random configurations using a generalization of
Newton-Raphson algorithm. Another way to sample configurations on a constrained
manifold consist in expressing some configuration variables with respect to
others~\cite{Cortes2002},\\ \cite{MirLam2018} when this is possible.\\ \cite{atlasRRT} propose another method based on non-linear projection. They cover the
constrained manifold by growing an atlas composed of local charts. This
approximation provides a probability distribution that is closer to the uniform
distribution over the manifold than the projection of a uniform distribution
over the configuration space.

@@

Manipulation planning is a particular instance of path planning were
some objects are moved by some robots. The configuration space of the
whole system is constrained by non-linear constrained due to the fact
that objects cannot move by themselves and should stay in a stable
pose when not grasped by a robot.  The accessible configuration space
is thus a union of sub-manifold as defined in the previous
section. Each of these manifold is moreover a foliation where each
leaf corresponds to a stable position of an object or to a grasp of an
object by a gripper. The geometrical structure of the problem is now
well understood~\cite{AlaSimLau89}. Although several instances of
manipulation problem exist like manipulation by
pushing~\cite{BenRiv1998}, or by throwing~\cite{WooZacLyn2017}, as well as
multicontact planning \cite{Bretl2006}, \cite{LenVaiYosKhe2013},
\cite{TonDelPetParManMan2018}, in this paper, we are only concerned with
prehensile manipulation planning.

@@

Rearrangement planning \cite{ota2004rearrangement}, \cite{LIS210}, consists in finding a sequence of manipulation paths that
move some objects from an initial pose to a final pose. The final configuration
of the robot is not specified. A simplifying asumption is the existence of a
monotone solution, that is a solution where each object is grasped once only and
is moved from its initial pose to its final pose \cite{stilman2008planning}, \cite{stilman2007manipulation}, \cite{srivastavaetal2014icra}, \cite{nieuwenhuisen2008effective}, \cite{ota2004rearrangement}. They mostly use two-level methods
composed of a symbolic or task planner and of a motion planner~\cite{lavalle2006planning}.

@@

@@

\noindent
{\small\texttt{xhost +\\docker run --rm -it \\ --name hpp --volume="/tmp/.X11-unix:/tmp/.X11-unix"\\ gitlab.laas.fr:4567/humanoid-path-planner/hpp-doc/\\ journal2020}}

and follow the instructions.

\subsection{Tutorial}

\subsection{Benchmark}

>>>>

\section{Introduction}

Robots in industrial manufacturing are today mostly programmed by
hand. They repeat the same motion thousands of times with a high
accuracy. Automating a task with some variability is however very
challenging since it requires more programming effort to integrate
sensors and some kind of motion planning in the process. The work
described in this paper is a small step to the simplification of
industrial process automation in the presence of some variability,
like the initial position of some object or the presence of unknown
obstacles. Although the work only covers the motion planning -- and
more accurately the manipulation planning -- issue, we think that it
is important not only to develop algorithms, but also to provide them
within an open-source software platform in order to make the
evaluation and then the integration of those algorithms easier.

This \paper therefore describes a software platform called \href{https://humanoid-path-planner.github.io/hpp-doc}{Humanoid Path
  Planner} taylored for manipulation planning in robotics. It can handle many types of robots, from manipulator arms to legged humanoid robots.
The main contributions are:
\begin{itemize}
\item an original and general modelling of prehensile manipulation based on
  non-linear constraints,
\item an original solver for non-linear constraints that is able to handle implicit and explicit constraints,
\item a manipulation planning algorithm that is able to tackle a great variety of manipulation planning problems,
\item an open-source software suite that implements all the above following state of the art development tools and methods.
\end{itemize}

This paper extends the work presented in \cite{mirabel-iros2016} and \cite{MirLam2018} with some new material.

The \paper is organized as follows. \Partie~\ref{sec:related-work} presents some related work for constrained motion planning and manipulation planning.
\Partie~\ref{partie:preliminary} introduces some preliminary notions like kinematic chains and Lie groups that are used to model the configuration space of each joint. \Partie~\ref{sec:constraints} introduces non linear constraints and solvers that are at the core of manipulation problem definition. \Partie~\ref{sec:manipulation-problem} defines the problem of prehensile manipulation in the general setting. \Partie~\ref{sec:manipulation-planning} provides a general algorithm that solves manipulation planning problems. Finally, \partie~\ref{sec:humanoid-path-planner} is devoted to the software platform implementing the notions introduced in the previous sections. Experimental results are provided for a large variety of problems.

Each \partie is implemented by one or several software packages. For some values that need to be computed, instead of providing formulas, we sometimes provide a link to the C++ or python implementation.

@@

Motion planning has given rise to a lot of research work for the past decades.
The problem consists in finding a collision-free path for a given system in
an environment populated with obstacles. The field covers a large variety of
different applications ranging from navigation for autonomous vehicles in
partially known environments~\cite{DarpaUrbanChallenge} to path planning for
deformable objects~\cite{LamKav2001}, \cite{RouFerTai2020} and many other applications like coverage path planning~\cite{coverage2001}, \cite{coverage2013}, or pursuit evasion planning~\cite{pursuit-evasion1999}.

Planning motions for high dimensional robots like humanoid robots or multi-arm
systems has been proven to be of a very high complexity~\cite{SchSha1983}, \cite{Canny1983}. Starting in the 1990's random sampling methods have been proposed
to solve the problem, trading the completeness property against efficiency in
solving problems in high dimensional configuration spaces \cite{KSLO1996}, \cite{HsuLatMot99}, \cite{KufLav00}. The latter methods are said probabistically complete since the probability to find a solution if one exists converges to 1 when the time of computation tends to infinity. Since then, asymptotically optimal random sampling algorithms have
been proposed~\cite{KarFra2011}.

@@

Some systems are subject to non-linear constraints. These non-linear
constraints define a sub-manifold of the configuration space the robot must
stay on. For instance, legged robots that must keep contact with the
ground and enforce quasi-static equilibrium, or multi-arm systems
grasping a same object are subject to this type of constraints. As the
volume of the constrained manifold is usually equal to zero, sampling
random configurations satisfying the constraints is an event of zero
probability. To sample configurations on the constrained manifold,
\cite{DalNakLamLau2009} and \cite{berenson2011} project random
configurations using a generalization of Newton-Raphson
algorithm. Another way to sample configurations on a constrained
manifold consists in expressing some configuration variables with
respect to others~\cite{Cortes2002}, \cite{MirLam2018} when this is
possible. \cite{atlasRRT} propose another method based on non-linear
projection. They cover the constrained manifold by growing an atlas
composed of local charts. This approximation provides a probability
distribution that is closer to the uniform distribution over the
manifold than the projection of a uniform distribution over the
configuration space. \cite{tb-rrt2016} propose a variation of the latter paper.
The main difference resides in the fact that the nodes
built on the tangent space are not immediately projected on the manifold.
\cite{KinMolKav2019} provide a rich overview
of the various approaches to motion planning with non-linear
constraints.

@@

Manipulation planning is a particular instance of path planning were
some objects are moved by some robots.
Although several instances of
manipulation problem exist like manipulation by
pushing~\cite{BenRiv1998}, or by throwing~\cite{WooZacLyn2017}, as well as
multicontact planning \cite{Bretl2006}, \cite{LenVaiYosKhe2013},
\cite{TonDelPetParManMan2018}, in this paper, we are only concerned with
prehensile manipulation planning.
The configuration space of the whole system is constrained by
non-linear constraints due to the fact that objects cannot move by
themselves and should stay in a stable pose when not grasped by a
robot.  The accessible configuration space is thus a union of
submanifolds as defined in the previous section. Each of these
manifolds is moreover a foliation where each leaf corresponds to a
stable position of an object or to a grasp of an object by a
gripper. The geometrical structure of the problem is now well
understood~\cite{AlaSimLau89}.

The first attempt to solve manipulation planning problems using random sampling
was proposed by \cite{simeon2004ijrr} where a reduction property reduces the
complexity of the problem.

@@

Rearrangement planning \cite{ota2004rearrangement}, \cite{LIS210},
\cite{KroBek2015}, \cite{LerPha2015} consists in finding a sequence of
manipulation paths that move some objects from an initial pose to a
final pose. The final configuration of the robot is not specified. A
simplifying asumption is the existence of a monotone solution, that is
a solution where each object is grasped once only and is moved from
its initial pose to its final pose \cite{stilman2008planning},
\cite{stilman2007manipulation}, \cite{srivastavaetal2014icra}, \cite{nieuwenhuisen2008effective},
\cite{ota2004rearrangement}. They mostly use two-level methods
composed of a symbolic or task planner and of a motion
planner~\cite{cambon:hal-01976081}.

Other contributions in manipulation planning explicitely address the problem of multi-arm manipulation \cite{GhaCorSim2009}, \cite{HarTsuLau2014}, \cite{DobBek2015}, \cite{XiaLerPha2017}.

\cite{schmitt19icra} propose an approach where two robots manipulate an object in a dynamic environment. The output of the algorithm is a sequence of controllers rather than a sequence of paths.

@@

Open-source software platforms are an important tool to make possible fair
comparison between algorithms.
Several software platforms are available for motion planning and/or manipulation
planning in the robotics community. The most popular one is undoubtedly OMPL \cite{ompl2012}. OMPL integrates many randomized path planning algorithms and is widely used for teaching. Recently, \cite{KinMolKav2019} proposed an extension for systems subject to non-linear constraints.

OpenRave~\cite{diankov_thesis} is
a software platform that addresses motion and manipulation planning. It includes
computation of forward kinematics.

One of the main differences between our solution and the previously cited ones resides in the way manipulation constraints are compiled into a graph. To our knowledge none of the previous solutions can handle a variety of problems as large as those exposed in Section~\ref{sec:benchmarks}

@@

\noindent
{\scriptsize\texttt{xhost +\\docker run --rm -it --name hpp \\--volume="/tmp/.X11-unix:/tmp/.X11-unix"\\ gitlab.laas.fr:4567/humanoid-path-planner/hpp-doc/\\ journal2020}}

\noindent
and follow the instructions.

\subsection{Experimental results}\label{sec:benchmarks}

In this section, we report on several experimental results obtained
with HPP software on constrained motion planning and on manipulation
planning problems. The raw data can be found in
\href{https://github.com/humanoid-path-planner/hpp_benchmark/tree/v4.10.0/2020-07-23}{\texttt{hpp\_benchmark}} package. We present here only a few test cases.
The benchmarks are run 20 times each on an Intel Core i7 at 2.60 GHz, with 32 Bbytes of RAM and 3 Gbytes of cache memory. For each test case, we report the minimum, maximum, mean and standard deviation of the time of computation on the one hand, and of the number of nodes in the roadmap built to solve the problem on the other hand.

\subsubsection{Constrained motion planning}

\begin{figure}
  \begin{center}
    \includegraphics[width=.49\linewidth]{figures/hrp2-q-init.png}
    \includegraphics[width=.49\linewidth]{figures/hrp2-q-goal.png}
  \end{center}
  \caption{Constrained motion planning for HRP-2 humanoid robot sliding on the ground in quasi-static equilibrium: the feet should stay horizontal with a fixed relative position and the center of mass should project between the feet. The initial configuration is shown on the left. The goal configuration is shown on the right. The algorithm a constrained RRT close to the one described in~\cite{DalNakLamLau2009}}.
  \label{fig:benchmark-hrp2}
\end{figure}
One test case concerns constrained motion planning. The robot is an HRP-2 humanoid robot in quasi-static equilibrium that can slide on the ground. This type of motion can be post-processed into a walking motion using the method described in \cite{dalibard:hal-00654175}. The results are displayed in Table~\ref{tab:benchmark-hrp2}.
\begin{table}
  \begin{center}
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    & min & max & mean & std dev \\
    \hline
    time (s) & 0.03 & 11.64 & 1.32 & 2.55 \\
    nodes & 4 &  136 & 32.40 & 30.72\\
    \hline
  \end{tabular}
  \end{center}
  \caption{Experimental results for HRP-2 sliding on the ground: time of computation and number of nodes.}
  \label{tab:benchmark-hrp2}
\end{table}

\subsubsection{Manipulation planning}

In this section, we present some experimental results of manipulation planning
problems obtained with M-RRT algorithm described in Section~\ref{sec:manipulation-planning}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figures/baxter-manipulation-boxes}
  \end{center}
  \caption{Manipulation problem with Baxter robot manipulating two small boxes. The robot is requested to swap the boxes.}
  \label{fig:baxter-manipulation-boxes}
\end{figure}

\begin{table}
  \begin{center}
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    & min & max & mean & std dev \\
    \hline
    time (s) & 0.84 & 15.60 & 7.60 & 4.48 \\
    nodes & 23 &  375 & 176.15 & 108.54 \\
    \hline
  \end{tabular}
  \end{center}
  \caption{Experimental results for Baxter robot manipulating two boxes on a table.}
  \label{tab:baxter-manipulation-boxes}
\end{table}

The first test case features robot Baxter manipulating two boxes on a table (see Figure~\ref{fig:baxter-manipulation-boxes}). The boxes are swaped between the initial and final configurations. The robot has two grippers and each box is equipped with a handle. The constraint graph thus contains 7 nodes. The experimental results are displayed in Table~\ref{tab:baxter-manipulation-boxes}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figures/pr2-manipulation-two-hands-init.png}
    \includegraphics[width=\linewidth]{figures/pr2-manipulation-two-hands-goal.png}    
  \end{center}
  \caption{Manipulation planning problem with PR-2 robot manipulating a box. The robot needs to flip the box upside down from an initial pose (top) to a goal pose (bottom).}
  \label{fig:pr2-manipulation-two-hands}
\end{figure}

\begin{table}
  \begin{center}
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    & min & max & mean & std dev \\
    \hline
    time (s) & 0.92 & 9.62 & 3.30 & 2.47 \\
    nodes & 6 &  111 & 32.90 & 31.63 \\
    \hline
  \end{tabular}
  \end{center}
  \caption{Experimental results for PR-2 robot manipulating a box on a table.}
  \label{tab:pr2-manipulation-two-hands}
\end{table}

The second test case features robot PR-2 manipulating a box on a table. The robot is resquested to flip the box upside down from an initial pose to a goal pose as represented in Figure~\ref{fig:pr2-manipulation-two-hands}. The robot is equipped with two grippers, the box is equipped with two handles. The constraint graph contains 7 nodes. Table~\ref{tab:pr2-manipulation-two-hands} shows the experimental results.

\begin{figure}
  \begin{center}
    \includegraphics[width=.49\linewidth]{figures/romeo-placard-init.png}
    \includegraphics[width=.49\linewidth]{figures/romeo-placard-goal.png}    
  \end{center}
  \caption{Manipulation planning problem with Romeo robot manipulating a placard. The robot needs to flip the placard by 180 degrees from an initial pose (left) to a goal pose (right).}
  \label{fig:romeo-placard}
\end{figure}

\begin{table}
  \begin{center}
  \begin{tabular}{|l|l|l|l|l|}
    \hline
    & min & max & mean & std dev \\
    \hline
    time (s) & 4.64 & 554.18 & 151.49 & 158.64 \\
    nodes & 27 &  2448 & 610.45 & 662.83 \\
    \hline
  \end{tabular}
  \end{center}
  \caption{Experimental results for Romeo robot manipulating a placard.}
  \label{tab:romeo-placard}
\end{table}

The third test case features humanoid robot Romeo manipulating a placard. The
robot is requested to rotate the placard by 180 degrees. The robot is equipped
with two grippers and the placard with two handles. Each handle is associated
to a unique gripper. The number of states of the contraint graph is thus 3.
